Laplacian envelope

### 2024-05-14 Notes/laplace.tex
changing planes

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{?.tex} with its contents
%       or include ?.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage[T1]{fontenc}
\usepackage{a4wide}
% \usepackage{stix2}
\usepackage[tables,graphics,draft,panels=paralist]{sl-preamble}
%\usepackage{amsldoc} for finding documentation only
\def\Snospace~{\S{}}
\renewcommand{\sectionautorefname}{\Snospace}
\renewcommand{\subsectionautorefname}{\Snospace}
\renewcommand{\subsubsectionautorefname}{\Snospace}
%
\graphicspath{{Figures/}}
%
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%================================================================================
% \makeatletter
% \DeclareFontEncoding{LS2}{}{\noaccents@}
% \makeatother

@@

The eigenvalue spectrum of \(\MM\) is closely related to its numerical range:
%
\begin{equation}\label{eq:numrange}
  \mathcal{W}(-\MM; \mathcal{V}) 
    = \set{\frac{\av{\mathbf{u}, -\MM \mathbf{u}}}{\av{\mathbf{u}, \mathbf{u}}}}
          {\mathbf{u} \in \mathcal{V}} 
    = \set{\frac{\nnrm{\boldsymbol{\Delta}\mathbf{u}}^2}{2\nrm{\mathbf{u}}^2}}
          {\mathbf{u} \in \mathcal{V}} .
\end{equation}
%
When \(\MM\) satisfies detailed balance (more generally when it commutes with its time reverse), the numerical range is the convex hull of the eigenvalues.

The other quantity of interest in \cref{eq:snrcoeffs} is the amplitude of the eigenmodes.
In analogy, we define the following functional (for real \(\mathbf{u}\))
%
\begin{equation*}
  \CI(\mathbf{u}) = \frac{(\eq \enc \mathbf{u}) (\mathbf{u}^\dag \w)}{r\mathbf{u}^\dag \mathbf{u}}
      = \frac{\sum_{i \neq j} \eqm_i \encm_{ij} (u_j - u_i) \sum_k \eqm_k u_k \wm_k} 
             {r\sum_{l} \eqm_l u_l^2} .
\end{equation*}
%
Following \cite{Lawler1988cheeger}, we can write
%
\begin{equation*}
  u_j - u_i = \intd[_{u_i}^{u_j}]{\alpha} 
      = \intd[_{-\infty}^{\infty}]{\alpha} 
        \brk{\vartheta(u_j - \alpha) \vartheta(\alpha - u_i) 
           - \vartheta(u_i - \alpha) \vartheta(\alpha - u_j)} ,
\end{equation*}
%
where \(\vartheta\) is the Heaviside step-function.
Then 
%
\begin{equation*}
  \begin{aligned}[t]
  \eq \enc \mathbf{u} &= \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
        \Fm\pot_{\CA_\alpha \CA_\alpha^c} - \Fm\dep_{\CA_\alpha \CA_\alpha^c} 
      - \Fm\pot_{\CA_\alpha^c \CA_\alpha} + \Fm\dep_{\CA_\alpha^c \CA_\alpha} }
  %
  \; \text{where} 
    \\
    &= \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
        \Fm_{\CA_\alpha \CA_\alpha^c} + \Fm_{\CA_\alpha^c \CA_\alpha} 
        - 2\Fm\dep_{\CA_\alpha \CA_\alpha^c} - 2\Fm\pot_{\CA_\alpha^c \CA_\alpha} } ,
  \end{aligned}
  %
  \begin{aligned}[t]
    \CA_\alpha &= \set{i}{u_i < \alpha} , 
    \\
    \Fm^\nu_{\CA \CB} &= \sum_{\substack{i \in \CA \\ j \in \CB}} 
                        \eqm_i \Wm^\nu_{ij} .
  \end{aligned}
\end{equation*}
%
Note that the first two terms of the integrand are of the form \( \Fm_{\CA \CB} = \Fm_{\CA \CB}\pot + \Fm_{\CA \CB}\dep \).
In exactly the same way as the proof of \cref{eq:initbnd}, we can argue that the two terms with the negative signs are indeed non-positive as they do not involve the negative diagonal elements of~\( \Fm_{ij}^\nu \).
Therefore the other terms provide an upper bound:
%
\begin{equation*}
\begin{aligned}
  \CI(\mathbf{u}) 
      &\leq \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
        \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
        \Fm_{\CA_\alpha \CA_\alpha^c} + \Fm_{\CA_\alpha^c \CA_\alpha} } 
  \\  &= \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
        \intd[_{-\infty}^{\infty}]{\alpha} \sum_{i \neq j} \Fm_{ij} \brk{
            \vartheta(u_j - \alpha) \vartheta(\alpha - u_i) 
          + \vartheta(u_i - \alpha) \vartheta(\alpha - u_j)
        }
  \\  &= \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
      \sum_{i \neq j} \Fm_{ij} \abs{u_j - u_i}
  \\  &= \frac{\av{\mathbf{u},\w} \aav{\abs{\boldsymbol{\Delta}\mathbf{u}}, \onem}}
              {r\nrm{\mathbf{u}}^2}.
\end{aligned}
\end{equation*}
%
Now we can apply the Cauchy-Schwartz inequality to both factors to find
%
\begin{equation*}
  \CI(\mathbf{u}) \leq \frac{\nrm{\w} \nnrm{\boldsymbol{\Delta}\mathbf{u}} \nnrm{\onem}}
                            {r\nrm{\mathbf{u}}} .
\end{equation*}
%
We also have \( \nrm{\w}^2 = \sum_i \eqm_i \wm_i^2 = 1 \).
Furthermore, using \cref{eq:constri}, \( \nnrm{\onem}^2 = \sum_{i \neq j} \eqm_i \frgm_{ij} \leq r \sum_i \eqm_i = r\).
Then, remembering \cref{eq:numrange},
%
\begin{equation}\label{eq:sqrtnumrange}
  \CI(\mathbf{u}) \leq \sqrt{\frac{2\av{\mathbf{u}, -\frg \mathbf{u}}}{r\av{\mathbf{u}, \mathbf{u}}}}
      = \sqrt{2\mathcal{W}(-\frg;\mathbf{u})/r}.
\end{equation}
%
Now, simply choosing \(\mathbf{u} = \evr_a \), one of the orthonormal eigenvectors, we find a bound on the eigenmodes:
%
\begin{equation}\label{eq:sqrtbound}
  \text{assuming} \quad 
  \eqm_i \frgm_{ij} = \eqm_j \frgm_{ji},
  \qquad \text{we have} \quad
  \CI_a \sqrt{r\tau_a} \leq \sqrt{2}.
\end{equation}
%
Numerical maximisation of the quantity above produces the same result as in \cref{sec:initial}:
the two-state model with deterministic transitions.
This model has \(\CI_1 = \tau_1 = 1\), suggesting that there's a factor of \(\sqrt{2}\) looseness in this bound.

@@

So we see that all four inequalities cannot be simultaneously saturated, and the bound must be loose.
Because we think that \cref{eq:sqrtbound} can be improved, we will rewrite it as
%
\begin{equation}\label{eq:gammabnd}
  \CI_a \sqrt{r\tau_a} \leq \sqrt{\cnst}.
\end{equation}
% 
We have shown that the smallest \(\cnst\) for which \cref{eq:gammabnd} is true is less than 2.
Considering the two-state model with deterministic transitions, we see that \(\cnst\) is at least 1.
For the \(M\)-state uniform serial model with deterministic transitions, in \cref{eq:unieig}, we find
%
\begin{equation*}
  \CI_n \sqrt{r\tau_n} = 
      \frac{2 \sqrt{2} \sin\frac{n\pi}{2} \cos\frac{n\pi}{2M}}
           {\prn{M \sin \frac{n\pi}{2M}}^2},
\end{equation*}
%
which is maximised at \(n=1, M \to \infty\).
Thus we see that \(\cnst \geq \frac{128}{\pi^4} = 1.31 \).

@@

Suppose there are two states, \(i,j\), with nonzero transitions from \(m\) for one \(\nu\).
Both must satisfy \cref{eq:oneexit} for the same \(\kktm^\nu_{mm}\), so
%
\begin{equation}\label{eq:twoexits}
  \Wm^\nu_{mi}, \Wm^\nu_{mj} > 0
  \qquad \implies \qquad
  \dgnm^\nu_{m,i} = \dgnm^\nu_{m,j}
  \quad \text{or}\quad
  \dgnm_{m,i} - \dgnm_{m,j} = - (-1)^\nu (\etwm_i - \etwm_j).
\end{equation}
%
Furthermore, for another state \(k\) with zero transition rate from \(m\) for the same \(\nu\):
%
\begin{equation}\label{eq:nonexit}
\begin{aligned}
  \Wm^\nu_{mi} > 0, \Wm^\nu_{mk} = 0 
  \;\;\; \implies \;\;\;
  \kktm^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i})
  \;\; \implies \;\;
  \dgnm^\nu_{m,k} &\geq \dgnm^\nu_{m,i} \\
  \text{or}\quad
  \dgnm_{m,k} - \dgnm_{m,i} &\geq - (-1)^\nu (\etwm_k - \etwm_i).
\end{aligned}
\end{equation}
%
If both \(\nu\) have nonzero rates from \(m\) to \(i\) and \(j\):
%
\begin{equation}\label{eq:twodoubleexits}
  \Wm\pot_{mi}, \Wm\dep_{mi}, \Wm\pot_{mj}, \Wm\dep_{mj} > 0
  \qquad \implies \qquad
  \dgnm_{m,i} = \dgnm_{m,j} 
  \quad \text{and} \quad
  \etwm_i = \etwm_j.
\end{equation}
%
The consequences of these inequalities are depicted in \cref{fig:planes}(\ref{fig:triangle_spacing}).
We see that \(n\) must be a global minimum of \(\dgn_m\), and if there are multiple transitions of a single type (potentiating or depressing) the targets are constrained to lie on a single line.
Even if there is no state \(n\) that is the target of both potentiating and depressing transitions, there will be two lines at \(45^\circ \) passing through any targets of potentiating/depressing transitions that demarcate the forbidden zone.

@@

\begin{figure}[ht]
\begin{center}
\begin{myenuma}
  \item \label{fig:triangle_spacing} \aligntop{ \hspace{-2em}%
  \includegraphics[width=0.4\linewidth]{spacing_triangle.svg}}
  \hspace{0.2em}
  \item \label{fig:scaling_plane}\aligntop{ \hspace{-1em}%
  \includegraphics[width=0.4\linewidth]{scaling_plane.svg}}
\end{myenuma}
\caption[Depicting constraints on topology]{Depicting constraints on topology.
(\ref*{fig:triangle_spacing}) The consequences of \cref{eq:oneexit,eq:doubleexit,eq:twoexits,eq:twodoubleexits,eq:nonexit} in the \( (\dgnm_{m,l},-\etwm_l) \)-plane, 
  with \(i\) a state with a nonzero depressing transition from \(m\), 
  \(j\) a state with a nonzero potentiating transition, 
  \(n\) a state with both depressing and potentiating transitions 
  and \(k\) a state with no nonzero transitions from \(m\).
  The positions of \(m\) and \(k\) are not yet constrained, other than being in the upper quadrant \wrt \(n\).
(\ref*{fig:scaling_plane}) Scaling of optimal models in the \((s,r)\)-plane.
  Optimal models on grey ``rays'' are related by scaling, and therefore have the same topology.
  The line at \(r = r_0\) is where we need the solutions.
  The line at \(s = \epsilon\) is close to where we can determine the topology.
\label{fig:planes}}
\end{center}
\end{figure}

@@

\begin{figure}[tb]
  \centering
  % Requires \usepackage{graphicx}
  \begin{myenuma}
    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_uni.svg}}\label{fig:serial_uni}
    \hspace{0.02\linewidth}
    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_sticky.svg}}\label{fig:serial_stk}
    \hspace{0.02\linewidth}
    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_shorten.svg}}\label{fig:serial_sht}
  \end{myenuma}
  \caption[Heuristic optimal models]{Heuristics for models that maximise memory curve.
  All unlabelled arrows have transition probabilities equal to 1.
  At one particular time, the model is (\ref{fig:serial_uni}) the uniform serial model with \(M\) states.
  At later times the model has (\ref{fig:serial_stk}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model.
  At earlier times, the model is (\ref{fig:serial_sht}) progressively shortened by reducing the transition probability to the end states.
  Once this reaches zero we discard the old end state and work on the new end state.
  } \label{fig:heuristicmodels}
\end{figure}

Looking at them models that were found by numerical maximisation in \cref{fig:envvid} 
(see the video for the full set of models) suggests we look at the particular set of models shown in \cref{fig:heuristicmodels}.

@@

These models' Laplace transforms can be computed analytically, as we will see below, leading to a heuristic envelope.
All of these calculations will be performed at \(f\potdep=\frac{1}{2}\), so the denominator will play no role.
We will work in units such that \(r=1\) temporarily, only restoring \(r\) at the end.
We study models with \(m\) strong and \(m\) weak states for a total of  \(M = 2 m\).

@@

This is only valid for
%
\begin{equation}\label{eq:univalid}
  2 \leq M_* \leq M,
  \qquad\implies\qquad
  2\sinh^2 \frac{y_*}{M} \leq s \leq 2\sinh^2 \frac{y_*}{2}.
\end{equation}
%
When \(M \gg 2\), the validity region is approximately
%
\begin{equation}\label{eq:univalidapprox}
  \frac{4.53327}{M^2} \lesssim s \lesssim 1.36423.
\end{equation}
%
When \(s \sim \CO(M^{-2}) \ll 1\), we have \(\beta \approx \pm \sqrt{2s}\) and the heuristic envelope is approximately
%
\begin{equation}\label{eq:unienvapprox}
  \hat{A}_*(s) \approx \frac{ 0.54203 }{ \sqrt{s} }.
\end{equation}
%
We will lok at a more accurate interpolation between different \(M\) in the \hyperref[sec:serial_shorten]{next section}.

@@

\subsection{Heuristic envelope}\label{sec:heuristicenv}

@@

\begin{figure}[tb]
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=0.8\linewidth]{LenvHeuristic.pdf}\\
  \caption[Heuristic envelope for the signal-to-noise ratio]
  {Heuristic envelope  for the signal-to-noise ratio, \cref{eq:henv}.
  Analytic expressions from \cref{eq:henvapprox} describe the yellow curve and are only valid to the right of each region.}\label{fig:heuristicenv}
\end{figure}

>>>>

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{?.tex} with its contents
%       or include ?.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage[T1]{fontenc}
\usepackage{a4wide}
% \usepackage{stix2}
\usepackage[tables,graphics,draft,panels=paralist,noeqcref]{sl-preamble}
%\usepackage{amsldoc} for finding documentation only
\def\Snospace~{\S{}}
\renewcommand{\sectionautorefname}{\Snospace}
\renewcommand{\subsectionautorefname}{\Snospace}
\renewcommand{\subsubsectionautorefname}{\Snospace}
%
\graphicspath{{Figures/}}
%
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%================================================================================
% \makeatletter
% \DeclareFontEncoding{LS2}{}{\noaccents@}
% \makeatother

@@

The eigenvalue spectrum of \(\MM\) is closely related to its numerical range:
%
\begin{equation}\label{eq:numrange}
  \mathcal{W}(-\MM; \mathcal{V}) 
    = \set{\frac{\av{\mathbf{u}, -\MM \mathbf{u}}}{\av{\mathbf{u}, \mathbf{u}}}}
          {\mathbf{u} \in \mathcal{V}} 
    = \set{\frac{\nnrm{\boldsymbol{\Delta}\mathbf{u}}^2}{2\nrm{\mathbf{u}}^2}}
          {\mathbf{u} \in \mathcal{V}} .
\end{equation}
%
When \(\MM\) satisfies detailed balance (more generally when it commutes with its time reverse), the numerical range is the convex hull of the eigenvalues.
For any vector \(\mathbf{u}\) that is perpendicular to \(\onev\), we define the following functional
%
\begin{equation*}
    \tau(\mathbf{u}) = - \frac{\mathbf{u}^\dag \mathbf{u}}{\mathbf{u}^\dag \frg \mathbf{u}}
      = \frac{1}{\mathcal{W}(-\frg, \mathbf{u})},
\end{equation*}
%
with the \(\mathbf{u}\) in \( \mathcal{W}(-\frg, \mathbf{u}) \) representing the vector space spanned by \(\mathbf{u}\).

The other quantity of interest in \cref{eq:snrcoeffs} is the amplitude of the eigenmodes.
In analogy, we define the following functionals (for real \(\mathbf{u}\))
%
\begin{equation*}
    \CI(\mathbf{u}) = \frac{(\eq \enc \mathbf{u}) (\mathbf{u}^\dag \w)}{r\mathbf{u}^\dag \mathbf{u}}
      = \frac{\sum_{i \neq j} \eqm_i \encm_{ij} (u_j - u_i) \sum_k \eqm_k u_k \wm_k} 
             {r\sum_{l} \eqm_l u_l^2} .
\end{equation*}
%
Following \cite{Lawler1988cheeger}, we can write
%
\begin{equation*}
  u_j - u_i = \intd[_{u_i}^{u_j}]{\alpha} 
      = \intd[_{-\infty}^{\infty}]{\alpha} 
        \brk{\vartheta(u_j - \alpha) \vartheta(\alpha - u_i) 
           - \vartheta(u_i - \alpha) \vartheta(\alpha - u_j)} ,
\end{equation*}
%
where \(\vartheta\) is the Heaviside step-function.
Then 
%
\begin{equation*}
  \begin{aligned}[t]
  \eq \enc \mathbf{u} &= \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
        \Fm\pot_{\CA_\alpha \CA_\alpha^c} - \Fm\dep_{\CA_\alpha \CA_\alpha^c} 
      - \Fm\pot_{\CA_\alpha^c \CA_\alpha} + \Fm\dep_{\CA_\alpha^c \CA_\alpha} }
  %
  \; \text{where} 
    \\
    &= \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
        \Fm_{\CA_\alpha \CA_\alpha^c} + \Fm_{\CA_\alpha^c \CA_\alpha} 
        - 2\Fm\dep_{\CA_\alpha \CA_\alpha^c} - 2\Fm\pot_{\CA_\alpha^c \CA_\alpha} } ,
  \end{aligned}
  %
  \begin{aligned}[t]
    \CA_\alpha &= \set{i}{u_i < \alpha} , 
    \\
    \Fm^\nu_{\CA \CB} &= \sum_{\substack{i \in \CA \\ j \in \CB}} 
                        \eqm_i \Wm^\nu_{ij} .
  \end{aligned}
\end{equation*}
%
Note that the first two terms of the integrand are of the form \( \Fm_{\CA \CB} = \Fm_{\CA \CB}\pot + \Fm_{\CA \CB}\dep \).
In exactly the same way as the proof of \cref{eq:initbnd}, we can argue that the two terms with the negative signs are indeed non-positive as they do not involve the negative diagonal elements of~\( \Fm_{ij}^\nu \).
Therefore the other terms provide an upper bound:
%
\begin{equation*}
\begin{aligned}
  \CI(\mathbf{u}) 
      &\leq \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
        \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
        \Fm_{\CA_\alpha \CA_\alpha^c} + \Fm_{\CA_\alpha^c \CA_\alpha} } 
  \\  &= \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
        \intd[_{-\infty}^{\infty}]{\alpha} \sum_{i \neq j} \Fm_{ij} \brk{
            \vartheta(u_j - \alpha) \vartheta(\alpha - u_i) 
          + \vartheta(u_i - \alpha) \vartheta(\alpha - u_j)
        }
  \\  &= \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
      \sum_{i \neq j} \Fm_{ij} \abs{u_j - u_i}
  \\  &= \frac{\av{\mathbf{u},\w} \aav{\abs{\boldsymbol{\Delta}\mathbf{u}}, \onem}}
              {r\nrm{\mathbf{u}}^2}.
\end{aligned}
\end{equation*}
%
Now we can apply the Cauchy-Schwartz inequality to both factors to find
%
\begin{equation*}
  \CI(\mathbf{u}) \leq \frac{\nrm{\w} \nnrm{\boldsymbol{\Delta}\mathbf{u}} \nnrm{\onem}}
                            {r\nrm{\mathbf{u}}} .
\end{equation*}
%
We also have \( \nrm{\w}^2 = \sum_i \eqm_i \wm_i^2 = 1 \).
Furthermore, using \cref{eq:constri}, \( \nnrm{\onem}^2 = \sum_{i \neq j} \eqm_i \frgm_{ij} \leq r \sum_i \eqm_i = r\).
Then, remembering \cref{eq:numrange},
%
\begin{equation}\label{eq:sqrtnumrange}
\begin{aligned}
  \CI(\mathbf{u}) &\leq \sqrt{\frac{2\av{\mathbf{u}, -\frg \mathbf{u}}}{r\av{\mathbf{u}, \mathbf{u}}}}
      = \sqrt{2\mathcal{W}(-\frg;\mathbf{u})/r} , &\quad \text{or} \quad
  \CI(\mathbf{u}) \sqrt{r \tau(\mathbf{u})} &\leq \sqrt{2}.
\end{aligned}
\end{equation}
%
Now, simply choosing \(\mathbf{u} = \evr_a \), one of the orthonormal eigenvectors, we find a bound on the eigenmodes:
%
\begin{equation}\label{eq:sqrtbound}
  \text{assuming} \quad 
  \eqm_i \frgm_{ij} = \eqm_j \frgm_{ji},
  \qquad \text{we have} \quad
  \CI_a \sqrt{r\tau_a} \leq \sqrt{2}.
\end{equation}
%
Numerical maximisation of the quantity above produces the same result as in \cref{sec:initial}:
the two-state model with deterministic transitions.
This model has \(\CI_1 = \tau_1 = 1\), suggesting that there's a factor of \(\sqrt{2}\) looseness in this bound, although we will see below that there is at least one model for which this quantity is higher.

@@

So we see that all four inequalities cannot be simultaneously saturated, and the bound must be loose.
Because we think that \cref{eq:sqrtbound} can be improved, we will rewrite it as
%
\begin{equation}\label{eq:gammabnd}
  \CI_a \sqrt{r\tau_a} \leq \sqrt{\cnst}.
\end{equation}
% 
We have shown that the smallest \(\cnst\) for which \cref{eq:gammabnd} is true is at most 2.
For the \(M\)-state uniform serial model with deterministic transitions, in \cref{eq:unieig}, we find
%
\begin{equation*}
  \CI_n \sqrt{r\tau_n} = 
      \frac{2 \sqrt{2} \sin\frac{n\pi}{2} \cos\frac{n\pi}{2M}}
           {\prn{M \sin \frac{n\pi}{2M}}^2},
\end{equation*}
%
which is maximised at \(n=1, M \to \infty\).
Thus we see that \(\cnst\) is at least \(\frac{128}{\pi^4} = 1.31 \).

@@

Suppose there are two states, \(i,j\), with nonzero transitions from \(m\) for one \(\nu\).
Both must satisfy \cref{eq:oneexit} for the same \(\kktm^\nu_{mm}\), so
%
\begin{equation}\label{eq:twoexits}
  \Wm^\nu_{mi}, \Wm^\nu_{mj} > 0
  \qquad \implies \qquad
  \dgnm^\nu_{m,i} = \dgnm^\nu_{m,j}
  \quad \text{or}\quad
  \dgnm_{m,i} - \dgnm_{m,j} = - (-1)^\nu (\etwm_i - \etwm_j).
\end{equation}
%
Furthermore, for another state \(k\) with zero transition rate from \(m\) for the same \(\nu\):
%
\begin{equation}\label{eq:nonexit}
\begin{aligned}
  \Wm^\nu_{mi} > 0, \Wm^\nu_{mk} = 0 
  \;\;\; \implies \;\;\;
  \kktm^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i})
  \;\; \implies \;\;
  \dgnm^\nu_{m,k} &\geq \dgnm^\nu_{m,i} \\
  \text{or}\quad
  \dgnm_{m,k} - \dgnm_{m,i} &\geq - (-1)^\nu (\etwm_k - \etwm_i).
\end{aligned}
\end{equation}
%
Therefore, the only nonzero transitions from \(m\) in \(\W^\nu\) must go to the global minimum of \(\dgn^\nu_m\).
This can also be seen by considering a perturbation that increases \(\Wm^\nu_{mi}\) and decreases \(\Wm^\nu_{mk}\) by the same amount (thus evading the diagonal constraint):
%
\begin{equation}\label{eq:exitminimum}
  \pdiff{\mathcal{L}}{\Wm^\nu_{mi}} - \pdiff{\mathcal{L}}{\Wm^\nu_{mk}}
    = \eqm_m \prn{\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i}} + \kktm^\nu_{mi} - \kktm^\nu_{mk}.
\end{equation}
%
If \( \dgnm^\nu_{m,k} > \dgnm^\nu_{m,i} \) we must have \( \kktm^\nu_{mk} > 0 \) and therefore \( \Wm^\nu_{mk} = 0 \).

If both \(\nu\) have nonzero rates from \(m\) to \(i\) and \(j\):
%
\begin{equation}\label{eq:twodoubleexits}
  \Wm\pot_{mi}, \Wm\dep_{mi}, \Wm\pot_{mj}, \Wm\dep_{mj} > 0
  \qquad \implies \qquad
  \dgnm_{m,i} = \dgnm_{m,j} 
  \quad \text{and} \quad
  \etwm_i = \etwm_j.
\end{equation}
%
The consequences of these inequalities are depicted in \cref{fig:planes}(\ref{fig:triangle_spacing}).
We see that \(n\) must be a global minimum of \(\dgn_m\), and if there are multiple transitions of a single type (potentiating or depressing) the targets are constrained to lie on a single line.
Even if there is no state \(n\) that is the target of both potentiating and depressing transitions, there will be two lines at \(45^\circ \) passing through any targets of potentiating/depressing transitions that demarcate the forbidden zone.

@@

\begin{figure}[ht]
\begin{center}
\begin{myenuma}
  \item \label{fig:triangle_spacing} \aligntop{ \hspace{-2em}%
  \includegraphics[width=0.4\linewidth]{spacing_triangle.svg}}
  \hspace{0.2em}
  \item \label{fig:scaling_plane}\aligntop{ \hspace{-1em}%
  \includegraphics[width=0.4\linewidth]{scaling_plane.svg}}
\end{myenuma}
\caption[Depicting constraints on topology]{Depicting constraints on topology.
(\ref*{fig:triangle_spacing}) The consequences of \cref{eq:oneexit,eq:doubleexit,eq:twoexits,eq:twodoubleexits,eq:nonexit} in the \( (-\etwm_l, \dgnm_{m,l}) \)-plane, 
  with \(i\) a state with a nonzero depressing transition from \(m\), 
  \(j\) a state with a nonzero potentiating transition, 
  \(n\) a state with both depressing and potentiating transitions 
  and \(k\) a state with no nonzero transitions from \(m\).
  The positions of \(m\) and \(k\) are not yet constrained, other than being in the upper quadrant \wrt \(n\).
(\ref*{fig:scaling_plane}) Scaling of optimal models in the \((s,r)\)-plane.
  Optimal models on grey ``rays'' are related by scaling, and therefore have the same topology.
  The line at \(r = r_0\) is where we need the solutions.
  The line at \(s = \epsilon\) is close to where we can determine the topology.
\label{fig:planes}}
\end{center}
\end{figure}

\subsubsection{Changing planes}\label{sec:planes}

Consider the setup of the previous section.
Suppose there was a nonzero transition in \(\W\pot\) from some state \(p\) to state \(m\).
Now \(m\) must be the minimum of \(\dgn\pot_p\), where \(j\) was the minimum of \(\dgn\pot_m\).

As we move from the \( (-\etwm_l, \dgnm_{m,l}) \)-plane in \cref{fig:triangle_spacing} to the \( (-\etwm_l, \dgnm_{p,l}) \)-plane, the different states will rise or fall depending on \(c_p - c_m\) and the value of \(\etwm_l\).
Let's focus on the rise/fall \emph{relative} to that of \(m\).
If \(c_p > c_m\), states to the left of \(m\) will rise and those to its right will fall relative to it.
If \(c_p < c_m\), states to the left of \(m\) will fall and those to its right will rise relative to it.

If state \(j\) is to the left of \(m\) we must have \(c_p > c_m\).
If \(j\) is to the right, as would be implied by the triangularity of \cref{sec:triangular} if it could be proven more generally,  we must have \(c_p < c_m\).

Similarly, suppose  there was a nonzero transition in \(\W\dep\) from some state \(q\) to state \(m\).
If state \(i\) is to the left of \(m\), as would be implied by the triangularity of \cref{sec:triangular}, we must have \(c_q > c_m\).
If \(i\) is to the right we must have \(c_q < c_m\).

The state \(j\) needs to do more than rise when we change planes.
It must rise further than the gap between the diagonals passing through \(j\) and \(m\), which is \( \frac{\kktm\pot_{mm}}{\eqm_m} \).
Regardless of the signs of the two factors, this rise is given by
%
\begin{equation}\label{eq:plane_rise}
  \prn{c_m - c_p} \prn{\etwm_m - \etwm_j} \geq \frac{\kktm\pot_{mm}}{\eqm_m}.
\end{equation}
%
A similar relation holds between \(c_q\), \(\etwm_i\) and \(\kktm\dep_{mm}\).

@@

\begin{figure}[tb]
  \centering
  % Requires \usepackage{graphicx}
  \begin{myenuma}
    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_uni.svg}}\label{fig:serial_uni}
    \hspace{0.02\linewidth}
    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_sticky.svg}}\label{fig:serial_stk}
    \hspace{0.02\linewidth}
    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_shorten.svg}}\label{fig:serial_sht}
  \end{myenuma}
  \caption[Semi-empirically optimal models]{Semi-empirical models that maximise memory curve.
  All unlabelled arrows have transition probabilities equal to 1.
  At one particular time, the model is (\ref{fig:serial_uni}) the uniform serial model with \(M\) states.
  At later times the model has (\ref{fig:serial_stk}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model.
  At earlier times, the model is (\ref{fig:serial_sht}) progressively shortened by reducing the transition probability to the end states.
  Once this reaches zero we discard the old end state and work on the new end state.
  } \label{fig:heuristicmodels}
\end{figure}

Looking at the models that were found by numerical maximisation in \cref{fig:envvid} 
(see the video for the full set of models) suggests we look at the particular set of models shown in \cref{fig:heuristicmodels}.
The two families of models are not identical, but are similar enough to provide some insight.

@@

These models' Laplace transforms can be computed analytically, as we will see below, leading to a semi-empirical envelope.
All of these calculations will be performed at \(f\potdep=\frac{1}{2}\), so the denominator will play no role.
We will work in units such that \(r=1\) temporarily, only restoring \(r\) at the end.
We study models with \(m\) strong and \(m\) weak states for a total of  \(M = 2 m\).

@@

This is only valid for
%
\begin{equation}\label{eq:univalid}
  2 \leq M_* \leq M,
  \qquad\implies\qquad
  2\sinh^2 \frac{y_*}{M} \leq s \leq 2\sinh^2 \frac{y_*}{2}.
\end{equation}
%
When \(M \gg 2\), the validity region is approximately
%
\begin{equation}\label{eq:univalidapprox}
  \frac{4.53327}{M^2} \lesssim s \lesssim 1.36423.
\end{equation}
%
When \(s \sim \CO(M^{-2}) \ll 1\), we have \(\beta \approx \pm \sqrt{2s}\) and the semi-empirical envelope is approximately
%
\begin{equation}\label{eq:unienvapprox}
  \hat{A}_*(s) \approx \frac{ 0.54203 }{ \sqrt{s} }.
\end{equation}
%
We will lok at a more accurate interpolation between different \(M\) in the \hyperref[sec:serial_shorten]{next section}.

@@

\subsection{Semi-empirical envelope}\label{sec:heuristicenv}

@@

\begin{figure}[tb]
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=0.8\linewidth]{LenvHeuristic.pdf}\\
  \caption[Semi-empirical envelope for the signal-to-noise ratio]
  {Semi-empirical envelope  for the signal-to-noise ratio, \cref{eq:henv}.
  Analytic expressions from \cref{eq:henvapprox} describe the yellow curve and are only valid to the right of each region.}\label{fig:heuristicenv}
\end{figure}
